\chapter{Duality Theory in Optimization}
\label{chap:04}

\paragraph{}In this section, we utilize the MC/MC framework to derive various duality results in convex and linear optimization. First we define what an optimization problem is:

\begin{defn}[Optimization Problem]\label{defn:040-opt-prob}
	An optimization problem is a problem of the form:
	\begin{align*}
		\text{Minimize}\quad   & f(x)                                   \\
		\text{subject to}\quad & x\in X\subset \mathbb{R}^m, g(x)\leq O
	\end{align*}
	where $f:\mathbb{R}^m\to \overline{\mathbb{R}}$ (resp. $g:\mathbb{R}^m\to \mathbb{R}^r,X\cap \operatorname{dom}(f),\{x\in X:g(x)\leq O\}$) is called the objective function (resp. constraint function, basic feasibility region, feasibility region) of the problem. The optimal value is written as $f^\ast$. The set of feasible solutions is written as $F$, with the set of optimal points written as $F^\ast$.
\end{defn}

\paragraph{}For convex problems, $f,X$ are convex, with $g$ having convex component functions, so $F$ is convex. In linear programming, $f,g$ are affine, with $X=\mathbb{R}^m$, so $F$ is polyhedral. For maximization problems, we regard it as the same as a minimization problem over $-f$ with other components fixed. For this note, we will always assume that the problem is convex. When dealing with equality constraints of the form $g(x)=O$ (often in this case $g$ is assumed to be affine), one can think as imposing inequality constraints $g(x)\leq O,-g(x)\leq O$.

\input{./src/04/041.tex}
\input{./src/04/042.tex}
\input{./src/04/043.tex}
\input{./src/04/044.tex}
\input{./src/04/045.tex}
\input{./src/04/046.tex}
